\documentclass[letterpaper, 12pt]{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{float}
\usepackage{enumitem}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{gensymb}
\usepackage{parskip}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{nicefrac}
\hypersetup{
    colorlinks=true,        % Enable colored links
    linkcolor=teal,         % Set color for internal links
    citecolor=teal,         % Set color for citations
    filecolor=teal,         % Set color for file links
    urlcolor=teal           % Set color for URLs
}

\DeclareMathOperator{\arcsec}{arcsec}

\newcommand{\example}[1]{\textcolor{blue}{\textbf{Example:} #1}}
\newcommand{\step}[2]{\textcolor{blue}{\textbf{Step #1:} #2}}

\begin{document}

\begin{center}
\textbf{{\Large AP Calculus BC Test 4}}
\end{center}

\subsection*{H23 Parametric equations}

A parametric equation is a pair of equations that define a curve in the plane. The equations are typically written in the form $x = f(t)$ and $y = g(t)$, where $t$ is a \textbf{parameter} that varies over some interval.

The curve defined by the parametric equations can be traced out by varying the parameter $t$ from its starting value to its ending value, \underline{drawing arrows} to indicate directionality.

\subsubsection*{Eliminating the parameter}
To eliminate the parameter, solve one of the equations for $t$ and then substitute that expression into the other equation.

\example{Given the parametric equations $x = 2t + 1$ and $y = t^2 - 3$, eliminate the parameter to find an equation in terms of $x$ and $y$.}

\begin{gather*}
x = 2t + 1 \\
t = \frac{x - 1}{2} \\
y = t^2 - 3 \\
y = \left(\frac{x - 1}{2}\right)^2 - 3 \\
\boxed{y = \frac{(x - 1)^2}{4} - 3}
\end{gather*}

For trigonometric functions, use a Pythagorean identity.

\example{Given the parametric equations $x = 4\cos(t)$ and $y = 4\sin(t)$, eliminate the parameter to find an equation in terms of $x$ and $y$. Determine the shape of the curve.}

\begin{gather*}
x = 4\cos(t) \\
y = 4\sin(t) \\
\end{gather*}

Using the Pythagorean identity $\sin^2(t) + \cos^2(t) = 1$:

\begin{gather*}
\left(\frac{x}{4}\right)^2 + \left(\frac{y}{4}\right)^2 = 1 \\
\boxed{\frac{x^2}{16} + \frac{y^2}{16} = 1}
\end{gather*}

The curve is a \fbox{circle}.

\subsubsection*{Tangent lines}

$$\frac{dy}{dx} = \frac{\frac{dy}{dt}}{\frac{dx}{dt}}$$
$$\frac{d^2y}{dx^2} = \frac{\frac{d}{dt}\left(\frac{dy}{dx}\right)}{\frac{dx}{dt}}$$

\example{Find the slope of the tangent line to the curve and the concavity of the curve defined by the parametric equations $x = 4cos(t)$ and $y = 4sin(t)$ at the point where $t = \frac{\pi}{4}$.}

\step{1}{Find the first derivative and plug in the value of $t$ to get the slope.}

\begin{gather*}
\frac{dy}{dx} = \frac{\frac{dy}{dt}}{\frac{dx}{dt}} \\
= \frac{4\cos(t)}{-4\sin(t)} \\
= -\cot(t) \\
-\cot(t)|_{t = \frac{\pi}{4}} = -1 \\
\boxed{m = 1}
\end{gather*}

\step{2}{Find the second derivative and plug in the value of $t$ to get the concavity.}

Since we know $\frac{dy}{dx} = -\cot(t)$, differentiate and divide by $\frac{dx}{dt}$ to find the second derivative:

\begin{gather*}
\frac{d^2y}{dx^2} = \frac{\frac{d}{dt}\left(-\cot(t)\right)}{\frac{dx}{dt}} \\
= \frac{\csc^2(t)}{-4\sin(t)} \\
-\frac{\csc^2(t)}{4\sin(t)}|_{t = \frac{\pi}{4}} = -\frac{\sqrt{2}}{2} \\
\boxed{\text{concave down}}
\end{gather*}

\subsubsection*{Horizontal and vertical tangents}

If $\frac{dy}{dt} = 0$ and $\frac{dx}{dt} \neq 0$, then there is a \textbf{horizontal tangent}.

If $\frac{dx}{dt} = 0$ and $\frac{dy}{dt} \neq 0$, then there is a \textbf{vertical tangent}.

\subsubsection*{Arc length formula}

For parametric equations $x = f(t)$ and $y = g(t)$, the arc length from $t = a$ to $t = b$ is given by the formula:

$$L = \int_a^b \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} dt$$

\end{document}